---
title: "Machine Learning Project"
author: "Toto Kamari"
date: "12/12/2020"
output: html_document
---

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har.

Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E).

Class A corresponds to the specified execution of the exercise, while the other 4 classes correspond to common mistakes. Participants were supervised by an experienced weight lifter to make sure the execution complied to the manner they were supposed to simulate. The exercises were performed by six male participants aged between 20-28 years, with little weight lifting experience. We made sure that all participants could easily simulate the mistakes in a safe and controlled manner by using a relatively light dumbbell (1.25kg).

## Data

```{r}
download.file('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv',destfile = './pml-training.csv', method = 'curl')

training <- read.csv('pml-training.csv')

download.file('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv',destfile = './pml-test.csv', method = 'curl')

testing <- read.csv('pml-test.csv')

library(ggplot2) 
library(caret)
```

The variable we are trying to predict is "classe".  Let's first start removing variables that don't have enough
variance.

```{r}
removevars <- nearZeroVar(training)
training <- training[,-removevars]
testing <- testing[,-removevars]
```

Let's now remove any variables that have significant number of NA's in them.

```{r}
training <- training[,colSums(is.na(training)) < .05*nrow(training)]
testing <- testing[,colSums(is.na(testing)) < .05*nrow(testing)]

training <- training[,-(1:6)]
testing <- testing[,-(1:6)]
```

## Running Prediction Models

Let's break the training data set into a training and validation set.

```{r}
intrain <- createDataPartition(y=training$classe, p=.75, list=FALSE)
training <- training[intrain,]
validation <- training[-intrain,]
```

Now let's start out with a random forest.

```{r}
set.seed(2184)

modfitone <- train(classe~., data=training,method="rf",trControl=trainControl(method="cv",number=5))
print(modfitone$finalModel)
```

And now we'll predict on the validation set.

```{r}
modfitone_prediction <- predict(modfitone, newdata=validation)
confusionMatrix(modfitone_prediction, factor(validation$classe))
```

Let us now use the boosting method.

```{r}
set.seed(2184)

modfittwo <- train(classe~., data=training,method="gbm",verbose=FALSE,
                   trControl=trainControl(method="repeatedcv",number=5))
print(modfittwo$finalModel)
```

And now we'll predict on the validation set.

```{r}
modfittwo_prediction <- predict(modfittwo, newdata=validation)
confusionMatrix(modfittwo_prediction, factor(validation$classe))
```

## Final Prediction

It appears we get better results using the random forest method.  So now we will use that on our test set.

```{r}
modfitfinal_prediction <- predict(modfitone, newdata=testing)
modfitfinal_prediction
```